Methods to prioritise pop-up active transport infrastructure

### 2020-05-09 article.Rmd
Remove words, add refs

The Covid-19 pandemic has transformed all sectors of the economy, not least transport.
[Demand for long distance trips has plummetted](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies have gone bust](https://arxiv.org/abs/2004.08460) [@iacus_estimating_2020; @jittrapirom_exploratory_2020].
Yet short distance travel patterns have also changed greatly and there has been an uptick in cycling in some places [@harrabin_boom_2020].
There are two main explanations for increased active transport during pandemic-related physical distancing measures: 1) with fewer opportunities to exercise, more people are taking to walking, running and cycling in their local area for mental and physical health; and 2) public transport options have been greatly reduced due to cuts in services and fear of being infected while in enclosed spaces, meaning that walking and particularly cycling may be the only way that workers can reach key destinations such as hospitals.
The second reason is particularly important given that many 'key workers' are low paid, with limited access to private automobiles.
From a physical activity perspective this change is welcome: obesity is a powerful predictor of all-cause mortality, including from Covid-19 [@docherty_features_2020].
From a physical distancing perspective, increased levels of walking and cycling,
the shift creates pressure on governments to ensure sufficient 'space for social distancing', while enabling increased active mobility for health and travel to key workplaces.
<!-- particularly in densly populated urban areas where pavements and cycleways may be narrow,  -->

@@

1. 'filtered permeability' interventions to prevent through traffic, with interventions as part of Salford's ['Liveable Streets' project](https://salfordliveablestreets.commonplace.is/) being a prominent example [@salfordcitycouncil_salford_2020]
<!-- (see [here](https://twitter.com/CatrionaSwanson/status/1258322956595453952) for a photo illustrating this type of intervention) -->
2. banning cars and to pedestrianise streets, greatly increasing the width of walkable space, New York's 'Open Streets' initiative being a prominent example [@litman_pandemicresilient_2020], and
3. the reallocation of one or more lanes of traffic to create pop-up cycleways, typically along arterial routes, with lanes reallocated using traffic cones in Auckland providing an early example @orsman_covid_2020.

@@

```{r t1, results='asis'}
# rj %>% 
#   st_drop_geometry() %>% 
#   select(name, highway2, maxspeed, pctgov, width) %>% 
#   table1::table1(~ highway2 + pctgov + width | maxspeed, data = ., )
```

>>>>

<!-- The Covid-19 pandemic has transformed all sectors of the economy, not least transport. -->
<!-- [Demand for long distance trips has plummetted](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies have gone bust](https://arxiv.org/abs/2004.08460) [@iacus_estimating_2020; @jittrapirom_exploratory_2020]. -->
Much attention has focused on the impacts of Covid-19 on long distance travel patterns [e.g. @iacus_estimating_2020; @jittrapirom_exploratory_2020].
Yet short distance travel patterns have also changed, with a notable uptick in active travel, particularly cycling, in some areas  [@harrabin_boom_2020].
The two main explanations for this are 1) the need increased need for exercise close to home during lockdowns for mental and physical health [@jimenez-pavon_physical_2020]; and 2) reduced public transport options and use [e.g. @tian_investigation_2020].

<!-- due to cuts in services and fear of being infected while in enclosed spaces, meaning that walking and particularly cycling may be the only way that workers can reach key destinations such as hospitals. -->

The second reason is particularly important given that many 'key workers' are low paid, with limited access to private automobiles.
From a physical activity perspective this change is welcome: obesity is a powerful predictor of all-cause mortality, including from Covid-19 [@docherty_features_2020].
From a physical distancing perspective, increased levels of walking and cycling,
the shift creates pressure on governments to ensure sufficient 'space for social distancing', while enabling increased active mobility for health and travel to key workplaces.
<!-- particularly in densly populated urban areas where pavements and cycleways may be narrow,  -->

@@

1. 'filtered permeability' interventions to prevent through traffic, with interventions as part of Salford's ['Liveable Streets' project](https://salfordliveablestreets.commonplace.is/) being a prominent example [@salfordcitycouncil_salford_2020]
<!-- (see [here](https://twitter.com/CatrionaSwanson/status/1258322956595453952) for a photo illustrating this type of intervention) -->
2. banning cars and to pedestrianise streets, greatly increasing the width of walkable space, New York's 'Open Streets' initiative being a prominent example [@litman_pandemicresilient_2020], and
3. the reallocation of one or more lanes of traffic to create pop-up cycleways, typically along arterial routes, with lanes reallocated using traffic cones in Auckland providing an early example [@orsman_covid_2020].

@@

```{r t1, results='asis'}
# rj %>%
#   st_drop_geometry() %>%
#   select(name, highway2, maxspeed, pctgov, width) %>%
#   table1::table1(~ highway2 + pctgov + width | maxspeed, data = ., )
knitr::include_graphics("https://user-images.githubusercontent.com/1825120/81471440-0aa1ba00-91e9-11ea-8e6e-dd9f7669bf98.png")
```
